### 2019-10-15 paper/v1.tex
Some suggested things to add + some light editing

\documentclass[prd,amsmath,amssymb,floatfix,superscriptaddress,nofootinbib,twocolumn]{revtex4-1}

\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\bea{\begin{eqnarray}}
\def\eea{\end{eqnarray}}
\newcommand{\vs}{\nonumber\\}
\newcommand{\Var}{\rm Var}
\newcommand{\vk}{\vec{k}}
\newcommand{\ec}[1]{Eq.~(\ref{eq:#1})}
\newcommand{\eec}[2]{Eqs.~(\ref{eq:#1}) and (\ref{eq:#2})}
\newcommand{\Ec}[1]{(\ref{eq:#1})}
\newcommand{\eql}[1]{\label{eq:#1}}
\newcommand{\rf}[1]{\ref{fig:#1}}
\newcommand{\sfig}[2]{
\includegraphics[width=#2]{../plots/#1}
        }
\newcommand{\sfigg}[2]{
\includegraphics[width=0.88\paperwidth]{../plots/#1}
        }
\newcommand{\sfigr}[2]{
\includegraphics[angle=270,origin=c,width=#2]{#1}
        }
\newcommand{\sfigra}[2]{
\includegraphics[angle=90,origin=c,width=#2]{#1}
        }
\newcommand{\Sfig}[2]{
   \begin{figure}[thbp]
   \begin{center}
    \sfig{../plots/#1.pdf}{\columnwidth}
    \caption{{\small #2}}
    \label{fig:#1}
     \end{center}
   \end{figure}
}
\newcommand{\Sfigg}[2]{
   \begin{figure*}[thbp]
    \sfigg{../plots/#1.pdf}{\columnwidth}
    \caption{{\small #2}}
    \label{fig:#1}
   \end{figure*}
}
\newcommand{\Spng}[2]{
   \begin{figure}[thbp]
   \begin{center}
    \sfigg{../plots/#1.png}{\columnwidth}
    \caption{{\small #2}}
    \label{fig:#1}
     \end{center}
   \end{figure}
}

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{simplewick}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{color}
\usepackage{enumitem}
\usepackage[linktocpage=true]{hyperref} 
\hypersetup{
    colorlinks=true,       
    linkcolor=red,         
    citecolor=blue,        
    filecolor=magenta,      
    urlcolor=blue           
}
\usepackage[all]{hypcap} 
\definecolor{darkgreen}{cmyk}{0.85,0.1,1.00,0} 
\definecolor{darkorange}{rgb}{1.0,0.2,0.0}
\newcommand{\scott}[1]{{\color{darkgreen} #1}}
\newcommand{\peikai}[1]{{\color{blue} #1}}
\newcommand{\prvs}[1]{{\color{magenta} #1}}
\newcommand{\AL}[1]{{\color{magenta} AL: #1}}
\newcommand{\MR}[1]{{\color{blue} MR: #1}}
\newcommand{\nl}{\\ \indent}
\newcommand{\RC}[1]{{\color{darkorange} #1}}
\begin{document}
\title{Large Scale Structure Reconstruction with Short-Wavelength Modes}
\author{\large Peikai Li}
\affiliation{Department of Physics, Carnegie Mellon University, Pittsburgh, PA 15213, USA}
\author{\large Scott Dodelson}
\affiliation{Department of Physics, Carnegie Mellon University, Pittsburgh, PA 15213, USA}
\author{\large Rupert A. C. Croft}
\affiliation{Department of Physics, Carnegie Mellon University, Pittsburgh, PA 15213, USA}
\affiliation{McWilliams Center for Cosmology, Carnegie Mellon University, Pittsburgh, PA 15213, USA}

\date{\today}
\begin{abstract}
  \noindent Large scale density modes cannot be measured directly through surveys but we can study them by measuring perturbations on small scales.
\RC{ Should really say instead something about  the fact that large scale modes are difficult to measure
    because they are sensitive to systematic observational  errors in
    galaxy surveys - not that they can't be measured}
  Cosmological perturbation theory predicts that second-order density inhomogeneities are a convolution of a short- and a long-wavelength mode. This arises physically because the growth of small scale structures varies depending on the large scale environment in which they reside. This induces small scale clustering in real space and an off-diagonal term in the two-point statistics in Fourier space. We can therefore build a quadratic estimator for the large scale field. We demonstrate that this quadratic estimator works well with an N-body simulation test \RC{what do we find? scales? + final
    speculative sentence
    about potential of method to do cosmology}.

\end{abstract}
\maketitle
\section{Introduction}
Measuring the distribution of structures on large scales has been one of the main goals of recent surveys despite it is hard to accomplish directly.
\RC{Should include brief examples of why it is very interesting - for example
  to constrain fNL, large scale anomalies, detect matter radiation-equality turnover - make it clear that a measurement would be exciting}
To tackle this problem, relations between long- and short-wavelength modes have been frequently discussed in recent years \cite{Baldauf:2011fer}\cite{Barreira:2017res}. The method of using the small scale position-dependent power spectrum to compute the squeezed-limit bispectrum also indicates that we can use small scale perturbation to infer large scale information to some extent \cite{Chiang:2014pos}\cite{Chiang:2015poss}. In order to fully achieve this goal, we develop in this work a generalized approach of building a quadratic estimator to measure long-wavelength modes indirectly. \\
Starting from second-order standard perturbation theory (SPT) \cite{Goroff:1986sts}\cite{Makino:1992fs}\cite{Jain:1994sop}, we can identify the second-order perturbation of a short-wavelength mode as a convolution of a short- and a long-wavelength mode. This can be further related to the cosmic microwave background (CMB) lensing case \cite{Hu:2002mr}, where the CMB temperature field has a second-order correction due to the gravitational field along line of sight. Similarly in our case the short-wavelength mode's nonlinear terms are related its large scale environment.
\\
The construction of a CMB lensing quadratic estimator makes use of the fact that small scale two-point correlations of CMB temperature modes have off-diagonal terms due to large scale perturbations caused by gravitational lensing. The same statistical feature shows up in our case as well -- the off-diagonal terms of the small scale correlations are no-longer zero, due to the effect of large scale modes. Thus we can create a quadratic estimator for long-wavelength modes using exactly the same formalism. \\
We begin with a brief review of SPT up to second-order in section \ref{sec1}. We then build the quadratic estimator and compute the detectability of the estimator in section \ref{sec2}. In section \ref{sec3} we present a concise study of the estimator with data from N-body simulation. Finally we conclude this work and discuss future work related to this in section \ref{sec4}. We use flat $\Lambda$CDM model with Planck Collaboration XVI (2014) \cite{Planck:2014cos} cosmological parameters in this work (same as parameters of the N-body simulation).

\section{Standard Perturbation Theory}\label{sec1}
In this section we review the SPT approach which describes small scale nonlinearity.
\RC{Say something about: see also a textbook, or the Bernardeau review reference - it looks to me that some of your notation is from the latter. Stating this explicitly could
be helpful for a reader.} 
We then compute the off-diagonal term of two short-wavelength modes in the squeezed limit. This non-zero off-diagonal term proportional to a long-wavelength mode will be the key factor of constructing the quadratic estimator. \\
Starting from a perfect pressureless fluid, the nonrelativistic cosmological fluid equations are continuity, Euler and Poisson equations:
\bea
\frac{\partial \delta(\vec{x},\tau)}{\partial \tau} &+&\vec{\nabla}\times [(1+\delta(\vec{x},\tau))\vec{v}(\vec{x},\tau)] =0\eql{fluid1} \\
(\frac{\partial}{\partial \tau} +\vec{v}(\vec{x},\tau)&\times&\vec{\nabla})\vec{v}(\vec{x},\tau)=-\frac{da}{d\tau}\frac{\vec{v}(\vec{x},\tau)}{a}-\vec{\nabla}\Phi \eql{eulereq}\\
&&\nabla^2 \Phi = 4\pi G a^2 \bar{\rho}_{\rm m} \delta(\vec{x},\tau) \eql{fluid3}
\eea
these equations fully determine the time evolution of the local density contrast $\delta$ and the peculiar velocity field $\vec{v}=d\vec{x}/d\tau$.\\
We can solve these equations in Fourier space with perturbative expansion \cite{Bernardeau:2002rev}:
\bea
{\delta}(\vk,\tau) &=&\sum_{n=1}^{\infty} {\delta}^{(n)}(\vk,\tau)=\sum_{n=1}^{\infty}D_1^{n}(\tau)\delta_{n}(\vk)  \eql{pert1}\\
{\theta}(\vk,\tau)&=&\sum_{n=1}^{\infty}{\theta}^{(n)}(\vk,\tau) \vs
&=&-\frac{d\,\ln D_1(\tau)}{d\tau}\sum_{n=1}^{\infty}D_1^{n}(\tau)\theta_{n}(\vk) \eql{pert2}
\eea
where $D_1$ is the linear growth factor \RC{define this e.g. $\delta(\tau)=D_{1}\delta(0)$}. The first order term ${\delta}^{(1)}$ corresponds to linear evolution. The linear power spectrum is given by this first order term via:
\be 
\langle {\delta^{(1)}}(\vk,\tau){\delta^{(1)}}(\vk',\tau) \rangle =(2\pi)^3 \delta_{\rm D}(\vk+\vk')P_{\rm lin}(k,\tau) \eql{lin}
\ee 
Substituting the perturbative series \ec{pert1} and \ec{pert2} into the Fourier transformed fluid equations \ec{fluid1}-\ec{fluid3}, we can solve for the second-order density contrast:
\be
{\delta}^{(2)}(\vk,\tau)=\int \frac{d^{3}\vk_{1}}{(2\pi)^3} F_2(\vk_1,\vk-\vk_1){\delta}^{(1)}(\vk_1,\tau) {\delta}^{(1)}(\vk-\vk_1,\tau) \eql{sorder} 
\ee
with 
\be
F_{2}(\vk_1,\vk_2)=\frac{5}{7}+\frac{2}{7}\frac{(\vk_1\cdot \vk_2)^2}{k_1^2 k_2^2}+\frac{\vk_1\cdot \vk_2}{2k_1k_2}(\frac{k_1}{k_2}+\frac{k_2}{k_1})\eql{f2}
\ee
Note that \ec{pert1}, \ec{pert2} and \ec{f2} are completely accurate only in an Einstein-de Sitter universe. Nonetheless for related calculations in a $\Lambda$CDM universe, the difference is found to be negligible \cite{Takahashi:2008to}, and thus we use the expressions from \ec{pert1} and \ec{f2} throughout this work.\\
We next compute the two-point correlations of two short-wavelength modes $\vk_s$ and $\vk_s'$, in the squeezed limit $\vk_l=\vk_s+\vk_s'$ \RC{why is this squeezed limit - is $\vk_s,\vk_s' \gg \vk_l$?}, up to second-order. Here $\vk_l$ corresponds to a long-wavelength mode, and $\tau$  is at the present day:
\bea 
&& \langle {\delta}(\vec{k}_s){\delta}(\vec{k}_s') \rangle|_{\vk_s+\vk_s'=\vk_l}
\vs 
&=&  \langle {\delta}^{(1)}(\vec{k}_s){\delta}^{(2)}(\vec{k}_s') \rangle+\langle {\delta}^{(2)}(\vec{k}_s){\delta}^{(1)}(\vec{k}_s') \rangle
\eea 
Substituting \ec{sorder} into the first bracket we get:
\bea 
\langle {\delta}^{(1)}(\vec{k}_s){\delta}^{(2)}(\vec{k}_s') \rangle =  \int \frac{d^3\vec{k}}{(2\pi)^3} F_2 (\vec{k},\vec{k}_s'-\vec{k})\vs
\times \langle {\delta}^{(1)}(\vec{k}_s){\delta}^{(1)}(\vec{k}_s'-\vec{k}){\delta}^{(1)}(\vec{k}) \rangle \eql{integral}
\eea 
\peikai{Need to rephrase (don't know how): We claim that generally correlation function of three Gaussian fields is non-zero under following consideration: if a three-point correlation function consists of two short-wavelength modes and one long-wavelength mode, we can do the following contraction:
\be 
\langle
\contraction{}{{\delta}}{{\delta^{(1)}\;\;\;\;}(}{}
{\delta^{(1)}}(\vk_{1}) {\delta^{(1)}}(\vk_{2}){\delta^{(1)}}(\vk_3) 
\rangle
=\langle {\delta^{(1)}}(\vk_{1}) {\delta^{(1)}}(\vk_{2}) \rangle {\delta^{(1)}}(\vk_3)  \eql{contraction}
\ee 
$\vk_1$ and $\vk_2$ corresponds to two arbitrary short modes and $\vk_3$ arbitrary long mode. The above equation is true since in a real life survey, we can only measure this long-wavelength mode very limited times. Thus it won't have any statistical property and we can safely extract the long-wavelength mode out of the bracket.}\\

\RC{Maybe this?: The RHS of \ec{integral} includes an integral over the
  three point
  correlation function of modes of a Gaussian field. If one of the modes has
  a  much longer wavelength than the others, then its contribution will be
  limited to a few examples, and
  a statistical average will not take place.  In this case, we can
  move the long
  wavelength mode outside the  bracket, so that the following is a
  good approximation:
\be 
\langle
\contraction{}{{\delta}}{{\delta^{(1)}\;\;\;\;}(}{}
{\delta^{(1)}}(\vk_{1}) {\delta^{(1)}}(\vk_{2}){\delta^{(1)}}(\vk_3) 
\rangle
=\langle {\delta^{(1)}}(\vk_{1}) {\delta^{(1)}}(\vk_{2}) \rangle {\delta^{(1)}}(\vk_3)  \eql{contraction},
\ee 
where $\vk_1$ and $\vk_2$ are two short modes and $\vk_3$ is the long
mode.
}

We can use this contraction twice in \ec{integral}. One use occurs when $|\vk| \ll |\vk_s|,\, |\vk_s'-\vk|$, when we can extract ${\delta}^{(1)}(\vk)$; the other is when $|\vk_s'-\vk| \ll |\vk|,\,|\vk_s|$.\\
Using \ec{lin}, \ec{integral} can be evaluated as:
\bea 
&&\int \frac{d^3\vec{k}}{(2\pi)^3} F_2 (\vec{k},\vec{k}_s'-\vec{k})\langle {\delta}^{(1)}(\vec{k}_s){\delta}^{(1)}(\vec{k}_s'-\vec{k}){\delta}^{(1)}(\vec{k}) \rangle \vs
&=&\int d^3\vk F_2(\vec{k},\vec{k}_s'-\vec{k}) \delta_{\rm D}(\vk_s+\vk_s'-\vk)P_{\rm 
lin}(k_s){\delta}^{(1)}(\vec{k}) \vs
&+&\int d^3\vk F_2(\vec{k},\vec{k}_s'-\vec{k}) \delta_{\rm D}(\vk_s+\vk)P_{\rm 
lin}(k_s){\delta}^{(1)}(\vk_s'-\vk) \vs
&=&2F_2(-\vk_s,\vk_s+\vk_s')P_{\rm lin}(k_s){\delta}^{(1)}(\vk_s+\vk_s')
\eea 
where we take advantage of the fact that $F_2$ is a symmetric function.
Finally we have:
\be 
\langle {\delta}(\vec{k}_s){\delta}(\vec{k}_s') \rangle =f(\vec{k}_s,\vec{k}_s'){\delta}^{(1)}(\vec{k}_l) \eql{2pt}
\ee 
with
\bea
f(\vec{k}_s,\vec{k}_s')&=&2F_2(-\vec{k}_s,\vec{k}_s+\vec{k}_s')P_{\rm lin}(k_s)\vs
&+&2F_2(-\vec{k}_s',\vec{k}_s+\vec{k}_s')P_{\rm lin}(k_s')       
\eea 
\ec{2pt} suggests that we can estimate long-wavelength modes using short-wavelength modes.
\RC{Move this sentence to \ref{sec3}: Also notice that second-order SPT gets less accurate higher in the nonlinear regime and will produce a bias of our quadratic estimator, we can see this effect in \ref{sec3}.}

\section{Quadratic Estimator}\label{sec2}
In this section we construct the quadratic estimator for long-wavelength modes starting from \ec{2pt}. We then forecast the detectability of the quadratic estimator for future surveys. As with the case of CMB lensing, we can write the general form of the estimator by averaging over pairs of short-wavelength modes:
\begin{eqnarray}
\hat{\delta}^{(1)}(\vec{k}_l)=A(\vec{k}_l)\int \frac{d^3 \vec{k}_s}{(2\pi)^3} g(\vec{k}_s,\vec{k}_s'){\delta}(\vec{k}_s){\delta}(\vec{k}_s') \eql{quadest},
\end{eqnarray} 
with $g$ being a weighting function, $\vk_s'=\vk_l-\vk_s$ and $A$  defined by requiring that $\langle \hat{\delta}^{(1)}(\vec{k}_l) \rangle={\delta}^{(1)}(\vec{k}_l)$:
\begin{eqnarray}
A(\vec{k}_l)=\bigg[\int \frac{d^3 \vec{k}_s}{(2\pi)^3} g(\vec{k}_s,\vec{k}_s')f(\vec{k}_s,\vec{k}_s')  \bigg]^{-1} \eql{a}
\end{eqnarray}
In the absence of shot noise the Gaussian noise is given by:
\be 
\langle \hat{\delta}^{(1)}(\vk_{l})\hat{\delta}^{(1)}(\vk_{l}') \rangle = (2\pi)^3 \delta_{\rm D}(\vk_{l}-\vk_{l}')[P_{\rm lin}(k_{l})+N(\vk_l)]
\ee 
with 
\begin{eqnarray}
&&N(\vec{k}_{l})=2A^2(\vk_{l})\vs
&&\times\int \frac{d^3 \vec{k}_{s}}{(2\pi)^3} g^2(\vec{k}_{s},\vk_l-\vec{k}_{s})P_{\rm nl}(k_{s})P_{\rm nl}(|\vk_l-\vk_s|)
\end{eqnarray}
where $P_{\rm nl}$ is the nonlinear power spectrum. Minimizing the noise term we can fix the form of $g$ to be:
\begin{eqnarray}
g(\vec{k}_{s},\vec{k}_{s}')=\frac{f(\vec{k}_{s},\vec{k}_{s}')}{2P_{\rm nl}(k_{s})P_{\rm nl}(k_{s}')}
\end{eqnarray} 
The noise term reduces simply to $N(\vk_l)=A(\vk_l)$. We find by testing that the value of $N$ is very insensitive to the choice of the lower limit of the integration \ec{a}, since most of the contribution comes from large $k_s$. We  discuss the choice of the upper limit later, in section \ref{sec3}.\\
Assuming Gaussian noise, the projected detectability of a $P(k_l)$ measurement using the quadratic estimator can be expressed as:
\be
\frac{1}{\sigma^{2}(k_l)}=\frac{V k_l^2 \Delta k }{(2\pi)^2}\bigg[\frac{P_{\rm lin}(k_l)}{P_{\rm lin}(k_l)+N(k_l) }\bigg]^2 \eql{error},
\ee
where $V$ is the volume of a survey and we compute the detectability for a set of narrow $k_l$-bins each separated by width $\Delta k$.\RC{ Write a couple of
  sentences about Figure 1, bearing in mind that you have put information in
  the caption already. Also remember to change the survey volume to
match DESI and state this.}

\Sfig{SN}{Long-wavelength power spectrum and its error from \ec{error} which can be expressed as $P(k_l)\sigma(k_l)$. We assume a toy survey of boxsize $L=6\, \rm Gpc$, thus volume $V=L^3$ and width $\Delta k = 2\pi/L$. Moreover, we set the integration range for $\vk_s$ from $0.02 \, \rm Mpc^{-1}$ to $0.15\,\rm Mpc^{-1}$. \RC{Put this in the main text: We will see in the next section \ref{sec3} that this choice of the upper limit of integration is very reasonable for our current construction at $z=0$.}}

\section{Demonstration with an N-Body Simulation}\label{sec3}
We test the power of the quadratic estimator using data from a cosmological
N-body simulation. We use the $z=0$ snapshot from BigMPDL, one of the MultiDark cosmological simulations \cite{Klypin:2014nov}. The cubical box side length of BigMDPL is $3.7\,\rm Gpc$. We use the dark matter particle data to compute
the matter density field, leaving the effect of using galaxies or halos to
trace the field \cite{Desjacques:2018rev} to  future work.\\
We use the code nbodykit \cite{Hand:2018nby} 
to measure the Fourier density modes, and \ec{quadest} to estimate the
long wavelength modes  from the measured short wavelength modes. 
How well the estimator works can be seen from Fig.~\rf{hist}, where we show
histograms of the ratio of the estimated mode amplitudes ($\hat{\delta}(\vk_l)$) to their true amplitudes ($\delta(\vk_l)$). The
two panels show the differences between a short wavelength mode cutoff of $k_s=0.15 \rm Mpc^{1}$
and  $k_s=0.25 \rm Mpc^{1}$ \RC{Comment on what the figure shows - for example individual mode amplitudes are unbiased when $k_s=
  0.15 \rm Mpc^{1}$, and predicted to within $\sim x \%$} \RC{moved this
  here from fig caption:
We can see that the results for $k_s=0.25 \rm Mpc^{1}$ are
biased (the mode is $\sim 20
\%$ too high)
due to the increasing inaccuracy of second-order SPT as we move into 
nonlinear scales}
\Sfig{hist}{Comparison of measured and predicted Fourier mode amplitudes: histograms of number counts of $|\hat{\delta}(\vk_l)/\delta(\vk_l)|$ for two different $k_s$ integration ranges.}
Another goal of this approach is to reproduce large scale structure in real space. To test this, we compute real space density fields by taking the
inverse transform
of both the directly measured long wavelength Fourier modes and
the long wavelength modes
predicted using the quadratic estimator. These density fields, $\delta$ and $\hat{\delta}$ respectively are plotted in Fig.~\rf{real}. We can see that...
\RC{describe how well it works based on what we can see in the figure}

\Sfig{real}{\RC{Sadly this figure can't really be two columns in a PRL - there isn't enough room. I think you should put it back to one column, but moving the color scale to the bottom or top would allow you to fill the column a bit better} Comparison of the true density field in
  the BigMPDL simulation ($\delta(\vec{x})$ computed using the
  directly measured large-scale modes, top row) and the density
  field from the quadratic estimator ($\hat{\delta}(\vec{x})$, middle row).
  The bottom row shows their difference. Each panel represents a slice through
  the simulation volume,  $3.7\,\rm Gpc$ wide, and
  one cell ($0.5\,\rm Gpc$) thick.  The upper limit of $\vk_s$ was $0.15 \, \rm Mpc^{-1}$. \RC{As discussed, remove this reference to
  S/N: We can compute the signal-to-noise ratio as: $S/N=\Var(\delta)/\Var(\hat{\delta}-\delta)\approx 3$. \peikai{comment?}}}

\section{Conclusion}\label{sec4}
In this paper, we have proposed a new and potentially powerful method to measure long-wavelength modes without having to actually measure large scale structure directly.
\RC{ State that this is good because we won't be sensitive to
  observational systematics, sky gradients, dust etc}
\RC{Give a short list of cool things we can do, with references: measure 3D clustering
  beyond the matter radiation equality turnover for the first time,
  look for evidence of general relativistic effects which are strongest
  on large scales (e.g., https://journals.aps.org/prd/abstract/10.1103/PhysRevD.85.023504 ), look for evidence of large scale anomalies, like in the CMB,
  constrain non-Gaussianity, 
  etc }

Extensions of our work suggest themselves and will be
relevant to attempts to apply it to observational data.
The noise term $N=A$ would reduce significantly if we go higher into the nonlinear regime while second-order SPT is no longer accurate in this case. Thus, to detect even larger scales, we may need to use contributions from higher order perturbations and expand the integration range to reduce the noise level. \\
Another issue relates to the range of long-wavelength modes that are detectable. In the N-body simulation,  $k_l$ for modes recovered ranged from
$1.7 \times 10^{-3} \, \rm Mpc^{-1}$ to $8.9\times 10^{-3} \, \rm Mpc^{-1}$.
If we try to measure smaller modes, the quadratic estimator approach would
become inaccurate since the contraction \ec{contraction} is no longer valid in this case.\\
Our current formalism for the quadratic estimator for long-wavelength modes is only valid in a box at a fixed time. As the length
scales are so large, a more realistic situation should deal with
observations on the light cone.
The expression for the quadratic estimator to be generalized to take this into account. Moreover, we notice that the upper limit of integration range should also increase if we consider higher redshifts, in order to include more nonlinear information. Finally, when applied to datasets consisting of
halos or galaxies, the approach should be modified to include galaxy bias. We
also leave this to future work.\\ 
\\ \\ \\ \\

\acknowledgements
\noindent We thank Adam Solomon, Duncan Campbell, Fabian Schmidt and Lam Hui for resourceful discussions.  SD, PL and RC were supported by U.S.\ Dept.\ of Energy contract DE-SC0019248. SD and RC also acknowledge support from NSF AST-1909193\\
The BigMDPL simulation was performed at LRZ Munich within the PRACE project pr86bu. The CosmoSim database (\url{www.cosmosim.org}) providing the file access is a service by the Leibniz-Institute for Astrophysics Potsdam (AIP).
%\input{main.bbl}
\bibliography{refs}
\end{document}